Stats for the convergence paper

### 2019-12-16 models/paper/convergence_paper_models.Rmd
new figs

```{r }
source('functions/grabdata.R')

@@

```{r }
# the frequentist fit reported in the paper
# watch out for factor levels
# lex typ: typical - random - reversed
# reg shift: nc - -40% - +40%
lme.fit2 = glmer(resp_post_reg ~ 1 + baseline_av_c * lex_typicality + pre_reg_av_c + reg_rate +
    (1 + baseline_av_c | participant_id) + 
    (1  | word), 
  data = test2, 
  family = binomial (link = logit),
  control=glmerControl(optimizer="bobyqa")
)
summary(lme.fit2)$coef %>% 
  xtable

@@

confint(lme.fit2, method = 'Wald', level = 0.95)
effects::allEffects(lme.fit2)
```

@@

t3la = t3 %>% 
  select(
    word,
    lex_typicality2,
    baseline.gcm,
    baseline.mgl,
    reg.av
  ) %>% 
  rename(
    GCM = baseline.gcm,
    MGL = baseline.mgl,
    participants = reg.av
  ) %>% 
  gather(type, baseline.averages, -word, -lex_typicality2)

t3lb = t3 %>% 
  select(
    word,
    lex_typicality2,
    individual.gcm,
    individual.mgl,
    post.reg.av
  ) %>% 
  rename(
    GCM = individual.gcm,
    MGL = individual.mgl,
    participants = post.reg.av
  ) %>% 
  gather(type, individual.averages, -word, -lex_typicality2)

@@

ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline average') +
  ylab('verb ESP average')
# ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
ggsave('fig4_2019.pdf', width = 12.5, height = 4)

>>>>

```{r }
# source('functions/grabdata.R')

celex = read_csv('data/celex_verbs_with_categories_csv_fixed_moder.txt')

celex = celex %>% 
  select(
    present.o,
    present,
    moder,
    regular
  ) %>% 
  rename(
    word = present.o,
    disc.old = present
  ) %>%
  mutate(
    category = case_when(
      regular == 'Y' ~ 'regular',
      regular == 'N' ~ 'irregular'
    )
  )  %>% 
  filter(!is.na(category))

# fixing transcription

transcription.matcher = read_csv('data/transcription_matcher.txt')

celex = left_join(celex, transcription.matcher)

celex[celex$disc == 'riN',]

# test set
baseline = read_csv('data/convergence_paper_baseline_data.txt')

baseline.verbs = read_csv('data/baseline_verbs.txt')
# esp data
espdata = read_csv('data/convergence_paper_esp_data.txt')

# esp only post-test with gcm weights
test2 = read_csv('data/convergence_paper_esp_test2_predictions.txt')

@@

```{r }
# the frequentist fit reported in the paper
# watch out for factor levels
# lex typ: typical - random - reversed
# reg rate: nc - -40% - +40%
test2$lexical_typicality = test2$lex_typicality
test2$regularisation_rate = ifelse(
  as.character(test2$reg_rate) == 'nc', 'no change',
  as.character(test2$reg_rate)
  )
test2$regularisation_rate = factor(test2$regularisation_rate, levels = c('no change', '-40%', '+40%'))

lme.fit2 = glmer(resp_post_reg ~ 1 + baseline_av_c * lexical_typicality + pre_reg_av_c + regularisation_rate +
    (1 + baseline_av_c | participant_id) + 
    (1  | word), 
  data = test2, 
  family = binomial (link = logit),
  control=glmerControl(optimizer="bobyqa")
)
summary(lme.fit2)$coef

plot(effects::effect('regularisation_rate', lme.fit2), main = '', xlab = 'regularisation shift', ylab = 'regular response in post-test')
plot(effects::effect('baseline_av_c:lexical_typicality', lme.fit2), multiline = T, lines=list(lty=c(1,2,3)), main = '', xlab = 'verb baseline mean', ylab = 'regular response in post-test')

@@

confint(lme.fit2, method = 'Wald', level = 0.95)
effects::allEffects(lme.fit2)

test2 %>% 
  mutate(
    reg_rate2 = factor(reg_rate, levels = c('-40%', 'nc', '+40%'))
  ) %>% 
  ggplot(aes(x = reg_rate2, y = post_reg_av, fill = reg_rate2)) +
  geom_jitter(aes(colour = reg_rate2), width = 0.3) +
    geom_violin(aes(alpha = 0.5)) +
    geom_boxplot(width = 0.2) +
    # stat_summary(fun.y=mean, geom="point", shape=16, size=4) +
    scale_fill_brewer(palette="Set2") +
    scale_colour_brewer(palette="Set2") +
    theme(
      text = element_text(size=30), 
      # axis.text.x = element_text(angle=90, vjust=0.5), 
      # axis.title.x=element_blank(), 
      legend.position = 'none'
      ) +
    xlab('bot regularisation shift') +
    ylab('participant post-test mean regularisation') +
    scale_x_discrete(labels=c('-40%', 'no change', '+40%'))
ggsave('reg_shift_av.pdf', width = 9, height = 6)

test2 %>% 
  group_by(
    word,
    baseline_av,
    lex_typicality
  ) %>% 
  summarise(
    mean = mean(resp_post_reg)
  ) %>% 
  ggplot(aes(x = baseline_av, y = mean, colour = lex_typicality, label = word)) +
  geom_label() +
  geom_smooth(method = 'lm', se = F) +
  scale_colour_brewer(palette="Dark2") +
    theme(
      text = element_text(size=30), 
      # axis.text.x = element_text(angle=90, vjust=0.5), 
      # axis.title.x=element_blank(), 
      legend.position = 'none'
      ) +
    xlab('verb baseline mean regularisation') +
    ylab('verb post-test mean regularisation')
ggsave('lex_typ_av.pdf', width = 9, height = 9)

sum1 = confint(lme.fit2, method = 'Wald') %>% data.frame
sum2 = summary(lme.fit2)$coef %>% data.frame
sum1$Term = rownames(sum1)
sum2$Term = rownames(sum2)
sum3 = left_join(sum2, sum1) %>% 
  rename(`Std error` = Std..Error,
    `z value` = z.value,
    `2.5%` = X2.5..,
    `97.5%` = X97.5..
    )
sum3 %>% dplyr::select(Term, Estimate, `Std error`, `z value`, `2.5%`, `97.5%`) %>% knitr::kable(digits = 2)

@@

t3la = t3 %>% 
  dplyr::select(
    word,
    lex_typicality2,
    baseline.gcm,
    baseline.mgl,
    reg.av
  ) %>% 
  rename(
    GCM = baseline.gcm,
    MGL = baseline.mgl,
    participants = reg.av
  ) %>% 
  gather(type, baseline.averages, -word, -lex_typicality2)

t3lb = t3 %>% 
  dplyr::select(
    word,
    lex_typicality2,
    individual.gcm,
    individual.mgl,
    post.reg.av
  ) %>% 
  rename(
    GCM = individual.gcm,
    MGL = individual.mgl,
    participants = post.reg.av
  ) %>% 
  gather(type, individual.averages, -word, -lex_typicality2)

@@

ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2, linetype = lex_typicality2)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
# ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
ggsave('fig4_2019.pdf', width = 12.5, height = 4)
